---
title: "Practice Problem : Time Series Analysis"
author: "Akkash K N R"
output: html_notebook
---

Congratulations on your new job! This time you are helping out Unicorn Investors with your data hacking skills. They are considering making an investment in a new form of transportation - JetRail. JetRail uses Jet propulsion technology to run rails and move people at a high speed! While JetRail has mastered the technology and they hold the patent for their product, the investment would only make sense, if they can get more than 1 Million monthly users with in next 18 months.
 
You need to help Unicorn ventures with the decision. They usually invest in B2C start-ups less than 4 years old looking for pre-series A funding. In order to help Unicorn Ventures in their decision, you need to forecast the traffic on JetRail for the next 7 months. You are provided with traffic data of JetRail since inception in the test file.

Evaluation Metric:

Root Mean Squre Error (RMSE) is the evaluation metric for this contest

Note: Public and private split is 20:80 where first 44 days in public and next 169 days in private

Link https://datahack.analyticsvidhya.com/contest/practice-problem-time-series-2/


### Load Library
```{r}
library(forecast)
library(xts)
library(h2o)
library(keras)
library(ggplot2)
library(quantmod)
library(zoo)
library(mxnet)
library(lubridate)
library(data.table)
library(caret)
library(xgboost)
```



### Load Dataset

```{r}
train <- read.csv('train.csv')
test <- read.csv('test.csv')
```

Lets explore the training Dataset
```{r}
summary(train)
str(train)
```

Lets create a copy of the dataset
```{r}
ntrain <- train
ntest <- test
```

Convert the DateTime from string to datetime
```{r}
ntrain$Datetime <- as.POSIXct(ntrain$Datetime ,format="%d-%m-%Y %H:%M")
ntest$Datetime <- as.POSIXct(ntest$Datetime ,format="%d-%m-%Y %H:%M")
```

```{r}
ntrain$istrain <- 1; ntest$istrain <- 0; ntest$Count <- NA
```


Generally a Time Series consist of 3 components:- Seasonality,Trend and Irreducable(Random) Error

In the below plot we can clearly understand the trend and seasonality of the dataset
```{r}
ndata <- ts(ntrain[,3],frequency = 24)
decomp <- decompose(ndata,type="additive")
```


```{r, fig.width=25, fig.height=15}
plot(decomp)
```


### Time Series Models

#### Implementing Holt's Winter Model

It is also know as three parameter smoothing technique. Level, Seasonality and Trend

$$
\begin{aligned}
 L_t &= \alpha(D_t/S_{t-s})+(1-\alpha)(L_{t-1}+T_{t-1}) \ &- (1) \\ 
 T_{t} &= \beta(L_t-L_{t-1}) +(1-\beta)T_{t-1}\ &- (2) \\
 S_{t} &= \gamma(D_t/L_{t}) +(1-\gamma)S_{t-s}\ &- (3) \\
 F_{t+m} &= (L_t+mT_{t})*S_{t+m-s} \ &-(4) \\
\end{aligned}
$$
Where L is Level, D is Preicted value, S is Seasonality, T is trend and m is the window size(how many periods ahead). Alpha, Beta and Gamma are coefficient parameters for Level, Trend and Seasonality. 

We need to change the alpha, beta, gamma value to attain the best score.

```{r}
pred_Holt <- HoltWinters(ndata)
pred_Holt
print(pred_Holt)
```

```{r}
plot(pred_Holt)
```

#### Implementing Forecast Exponential Smoothening

Exponential Smoothening is suitable for forecasting data with no trend or seasonal pattern.

$$
\begin{aligned}
 F_{t+1} &= \alpha*D_t+(1-\alpha)F_t \\
\end{aligned}
$$
F(t+1) is the future forecast, alpha is coefficient parameter Dt is the output at time t, F(t) is the mean forecast. of time "t"

```{r}
#ets
pred_ets <- ets(ndata)
pred_ets
```

```{r}
plot(forecast(pred_ets))
```

#### Implementing Auto ARIMA

ARIMA is a "Auto Regression Integration with Moving Average". ARIMA(p,d,q), where "p" is the order of auto regression, "q" is the order of moving average and "d" is first differencing involved.

```{r}
#ARIMA
pred_ar <- auto.arima(diff(ndata))
pred_ar
```

#### Implementing STL

Also known as "Seasonal & Trend decomposition using Loess" is a type od decomposition, where Loess is a method for estimating nonlinear relationship.


```{r}
#STL
pred_STL <- stlm(ndata,modelfunction = ar ,s.window = 24)
```

Now lets predict the future...

```{r}
fcst_Holt <- forecast(pred_Holt,h=5112)

fcst_ets <- forecast(pred_ets,h=5112)

fcst_ar <- forecast(pred_ar,h=5112)

fcst_stl <- forecast(pred_STL,h=5112)
```

Plot the prediction
```{r}
plot(fcst_Holt)
```
```{r}
plot(fcst_ets)
```
```{r}
plot(fcst_ar)
```
```{r}
plot(fcst_stl)
```

Of all the models ETS gave the best results in the public test split so far.

Now lets implement more advance models.



#### Implementing Neural Network,

Artificial Neural Network are used to find the nonlinear relationship. 

http://neuralnetworksanddeeplearning.com/chap1.html

$$
\begin{aligned}
 N_j &= b_j+\sum_{j=1}^{n}w_{i,j}x_i \\ 
\end{aligned}
$$

```{r}
nnet <- nnetar(diff(ndata))
```


```{r}
plot(forecast(nnet))
```

```{r}
fcst_nnte <- forecast(nnet, h= 5112)
```

```{r}
plot(fcst_nnte)
```


```{r}
submission <- data.frame('ID'=test$ID,'Count'=fcst_Holt)
colnames(submission) <- c('ID','Count')
filename <- paste('ak_hw_',format(Sys.time(),"%Y%m%d%H%M%s"),sep = '_')
write.csv(submission,paste0(filename,'.csv',collapse = ''),row.names = FALSE)
```